---
title: "A Monopoly Simulation"
output: html_notebook
---

In this session, we will simulate simplified games of **Monopoly**.

### Moving around the board

Basically, a Monopoly board has 40 spaces - Players take it in turns to roll two dice and traverse around the  board according to the sum of the dice values.

**Simulate turns of a Single player**
```{r}
# Number of turns to take
num_turns <- 100000 

# Start on the GO space
current_board_position <- 0 

positions_visited <- rep(0, num_turns) # Store the positions

# Use a for loop
# to simulate a number of turns
for (turn in 1:num_turns) {

  # Roll two dice
  die_values <- sample(c(1:6), 2, replace = TRUE)

  ## Move player position
  
  # Number of positions to move
  plus_move <- sum(die_values)
  # Compute new board position
  new_board_position <- current_board_position + plus_move

  # Update board position (this corrects for the fact the board is circular)
  current_board_position <- (new_board_position %% 40)

  # Store position visited
  positions_visited[turn] <- current_board_position

}
```

By increasing the number of turns taken, what distribution does the set of simulated board positions converge towards?

```{r}
hist(positions_visited, breaks = seq(0, 40, len = 41), right = FALSE)
```

### Going to Jail

If a player lands on to Go To Jail space they must move immediately to the Jail space. Extend your code to include the possibility of going to jail. Here, assume that once in jail, the player continues as normal on the next turn.

```{r}
# Number of turns to take
num_turns <- 100000 

# Start on the GO space
current_board_position <- 0 
# The go to jail space
go_to_jail_position <- 30 
# The Jail space
jail_position <- 10 

positions_visited <- rep(0, num_turns) # Store the positions

# Use a for loop
# to simulate a number of turns
for (turn in 1:num_turns) {

  # Roll two dice
  die_values <- sample(c(1:6), 2, replace = TRUE)

  ## Move player position

  # Number of positions to move
  plus_move <- sum(die_values)
  # Compute new board position
  new_board_position <- current_board_position + plus_move

  # if land on GO TO JAIL square
  # then go backwards to the JAIL square
  if (new_board_position == go_to_jail_position) {
    new_board_position <- jail_position
  }

  # Update board position (this corrects for the fact the board is circular)
  current_board_position <- (new_board_position %% 40)

  # Store position visited
  positions_visited[turn] <- current_board_position

}
```

```{r}
hist(positions_visited, breaks = seq(0, 40, len = 41), right = FALSE)
```
You can also go to jail, if you roll three doubles (both dice having the same value) in a row. Update your code to allow for the possibility of going to Jail with three doubles.

```{r}
# Number of turns to take
num_turns <- 100000 

# Start on the GO space
current_board_position <- 0 
# The go to jail space
go_to_jail_position <- 30 
# The Jail space
jail_position <- 10 

positions_visited <- rep(0, num_turns) # Store the positions

# Use a for loop
# to simulate a number of turns
for (turn in 1:num_turns) {
  
  # Set double counter to zero
  double_counter <- 0
  
  # Roll 1-3(Max) times
  for (j in 1:3){

    # Roll two dice
    die_values <- sample(c(1:6), 2, replace = TRUE)

    # if we have rolled a double for the third time, we proceed straight to jail
    if ((die_values[1] == die_values[2]) & (double_counter == 2 )) {
      current_board_position <- jail_position
      break
    }
    
    ##Otherwise -  Move player position

    # Number of positions to move
    plus_move <- sum(die_values)
    # Compute new board position
    new_board_position <- current_board_position + plus_move

    # if land on GO TO JAIL square
    # then go backwards to the JAIL square
    if (new_board_position == go_to_jail_position) {
      new_board_position <- jail_position
    }

    # Update board position (this corrects for the fact the board is circular)
    current_board_position <- (new_board_position %% 40)

    # break out of loop if we roll a non-double
    if (die_values[1] != die_values[2]) {
      break
    } else { # increment double counter
      double_counter <- double_counter + 1
    }
  }
  
  # store final position visited
  positions_visited[turn] <- current_board_position

}
```

```{r}
hist(positions_visited, breaks = seq(0, 40, len = 41), right = FALSE)
```


### Further Exercise

Now consider building a more complex Monopoly simulation by incorporating more complex aspects of the game - the purchase of properties, a ledger for each player and chance and community cards.

```{r}
# Number of games to play
num_games <- 1000  
# Number of turns to take
num_turns <- 1000  

# Start on the GO space
current_board_position <- 0 
# The go to jail space
go_to_jail_position <- 30 
# The Jail space
jail_position <- 10 

# Vector of squares containing properties
properties_that_can_be_bought <- c(1, 3, 5, 6, 8, 9, 11, 12, 13, 14, 15, 16, 18, 19, 
                                   21, 23, 24, 25, 26, 27, 28, 29, 31, 32, 34, 35, 37, 39)

# Vector to store number of turns to buy all properties
time_to_buy_all_properties <- rep(0, num_games)

# Simulate multiple games
for (game in 1:num_games) {

  positions_visited <- rep(0, num_turns)
  positions_purchased <- rep(0, 40)
  properties_bought <- rep(0, num_turns)

  # Use a for loop 
  # to simulate a number of turns
  for (turn in 1:num_turns) {

    # Roll two dice
    die_values <- sample(c(1:6), 2, replace = TRUE)

    ## Move player position

    # Number of positions to move
    plus_move <- sum(die_values)
    # Compute new board position
    new_board_position <- current_board_position + plus_move


    # if land on GO TO JAIL square
    # then go backwards to the JAIL square
    if (new_board_position == go_to_jail_position) {
      new_board_position <- jail_position
    }

    # Update board position (this corrects for the fact the board is circular)
    current_board_position <- (new_board_position %% 40)

    # if we can on a square that can be purchased and which has not been
    # purchased (note R uses 1-indexing for arrays)
    if (positions_purchased[current_board_position+1] == 0) {
      if (current_board_position %in% properties_that_can_be_bought) {
        positions_purchased[current_board_position + 1] <- 1
      }
    }

    # Store position visited
    positions_visited[turn] <- current_board_position

    # Store number of properties bought
    properties_bought[turn] <- sum(positions_purchased)

    # Check if all properties are gone
    if (properties_bought[turn] == length(properties_that_can_be_bought)) {
      time_to_buy_all_properties[game] <- turn
      break
    }
  }

}
```

```{r}
hist(time_to_buy_all_properties, breaks = 20)
```
















